    
```
```

### Billboard Hot100 2014 in brief


Hossein Parsa
```
```

#### Introduction

The [Billboard Hot100](http://www.billboard.com/charts/hot-100) is a chart that ranks the best performing singles of the United States and is published weekly by Billboard Magazine. The ranking is based on radio play, online streaming and sales figures.

#### Data acquisition and processing

The data for the Billboard Hot100 2014 chart was scraped using `Python` code which is available at [GitHub](https://github.com/hparsa741/Data) in two steps of downloading source code and extracting the data. Once the data is available in csv format we start analyzing in `R` by reading the file.


```{r, echo=TRUE, cache=FALSE, message=FALSE, warning=FALSE}
data <- data.frame(read.csv("Hot100_2014.csv"))
library(sqldf)
library(dplyr)
library(ggplot2)
```

#### Uniques
  
Number of Unique Artists
  
```{r, echo=TRUE, cache=TRUE}
length(unique(data$singer))
```
  
Number of unique Songs
  
```{r, echo=TRUE, cache=TRUE}
dim(unique(data[c("singer", "song")]))[1]
```
  
Average song per artist
  
```{r, echo=TRUE, cache=TRUE}
dim(unique(data[c("singer", "song")]))[1]/length(unique(data$singer))
```
  
#### Position 1
  
Artists and count of weeks holding Position 1 
  
```{r, echo=TRUE, cache=TRUE, message=FALSE, warning=FALSE}
pos1<-data[which(data$pos == 1),]
arrange(summarize(group_by(pos1, singer), weeks = n()), desc(weeks))
```
  
Songs and count of weeks holding Position 1 
  
```{r, echo=TRUE, cache=TRUE}
pos1song<-aggregate(pos1$pos, list(singer=pos1$singer, song=pos1$song), sum)
names(pos1song)<-c("singer", "song", "weeks")
arrange(pos1song, desc(weeks))
```
  
#### Songs weeks on chart distribution
  
```{r, echo=TRUE, cache=TRUE}
sscount<-summarize(group_by(data, singer, song), count = n())
ggplot(sscount, aes(count))+geom_histogram(aes(y = ..density.., fill= ..count..), binwidth=1)+geom_density(color='red')+scale_fill_gradient("Count", low = "lightgreen", high = "darkgrey")+labs(title = "Weeks on chart distribution - song")+xlab("Weeks on chart")+guides(fill = guide_legend(title = "count of \nsongs", title.position = "top"))+theme(panel.background = element_rect(fill = "white"))
```
  
#### Artists with 5 or more unique song
  
```{r, echo=TRUE, cache=TRUE, message=FALSE, warning=FALSE}
sqldf("with t as (select distinct singer, song from data) select singer, count(0) as countOfSong from t group by singer having count(0) > 4 order by 2 desc")
```
  
#### Songs per artist distribution
  
```{r, echo=TRUE, cache=TRUE, message=FALSE, warning=FALSE}
ggplot(sqldf("with t as (select distinct singer, song from data) select singer, count(0) as countOfSong from t group by singer order by 2 desc"), aes(countOfSong))+geom_histogram(aes( fill= ..count..), binwidth=1)+scale_fill_gradient("Count", low = "darkgreen", high = "darkgrey")+labs(title = "Song per artist distribution")+xlab("Song per artist")+ylab("")+guides(fill = guide_legend(title = "count of \nartist", title.position = "top"))+theme(panel.background = element_rect(fill = "white"))
```
  
#### Top weeks on chart artists
  
45 or more weeks on chart
  
```{r, echo=TRUE, cache=TRUE}
sqldf("select singer, count(0) as Weeks_on_Chart from (select distinct singer, weekNo from data) group by singer having count(0) >= 45 order by 2 desc, 1")
```
  
Weeks on chart distribution
  
```{r, echo=TRUE, cache=TRUE}
ggplot(sqldf("select singer, count(0) as Weeks_on_Chart from (select distinct singer, weekNo from data) group by singer order by 2 desc, 1"), aes(Weeks_on_Chart))+geom_histogram(aes(y=..density.., fill= ..count..), binwidth=1)+geom_density(color='red')+scale_fill_gradient("Count", low = "darkgreen", high = "darkgrey")+labs(title = "Weeks on Chart Artist distribution")+xlab("Weeks on Chart")+ylab("")+guides(fill = guide_legend(title = "count of \nartist", title.position = "top"))+theme(panel.background = element_rect(fill = "white"))
```
    
Top 15 Artists appeared most on chart
  
```{r, echo=TRUE, cache=TRUE}
singerOnChart<-aggregate( pos ~ singer, data = data, FUN = length)
names(singerOnChart)<-c("Artist", "Weeks")
head(singerOnChart[with(singerOnChart, order(-Weeks)),], 15)
```
  
#### Top weeks on chart songs
  
30 or more weeks on chart
  
```{r, echo=TRUE, cache=TRUE}
sqldf("select singer, song, count(0) as Weeks_on_Chart from (select distinct singer, song, weekNo from data) group by singer, song having count(0) >= 30 order by 3 desc, 1")
```
  
Weeks on chart distribution
  
```{r, echo=TRUE, cache=TRUE}
ggplot(sqldf("select singer, song, count(0) as Weeks_on_Chart from (select distinct singer, song, weekNo from data) group by singer, song order by 3 desc, 1"), aes(Weeks_on_Chart))+geom_histogram(aes(y=..density.., fill= ..count..), binwidth=1)+geom_density(color='red')+scale_fill_gradient("Count", low = "darkgreen", high = "darkgrey")+labs(title = "Weeks on Chart Song distribution")+xlab("Weeks on Chart")+ylab("")+guides(fill = guide_legend(title = "count of \nartist", title.position = "top"))+theme(panel.background = element_rect(fill = "white"))
```
  
Positions of top 10 longest on chart songs
    
```{r}
topSong<-aggregate(data$pos, list(singer=data$singer, song=data$song), length)
head(topSong[order(-topSong[,3]),],10)
topSong<-head(topSong[order(-topSong[,3]),],10)
topAll <- data[data$song %in% head(topSong[order(-topSong[,3]),],10)$song & data$singer %in% head(topSong[order(-topSong[,3]),],10)$singer,]

ggplot(topAll, aes(weekNo, pos))+geom_line(aes(col=topAll$song),size=1)+labs(title = "Position of longest on chart songs", colour="Song")+xlab("Week Number")+ylab("")+guides(fill = guide_legend(title = "count of \nartist", title.position = "top"))+theme(panel.background = element_rect(fill = "white"))
```
  
```{r, echo=FALSE, cache=TRUE, eval=FALSE}
#### Sons from different artists having exactly same titles
sqldf("with tout as (select distinct data.singer, data.song from data, (select distinct song, count(0) from (select singer, song from data group by singer, song) group by song having count(0) >1 order by 2) t where t.song=data.song) select song, GROUP_CONCAT(singer, ', ') as 'List of Artists' , count(0) from tout group by song")
```
  
  
```
```